### 2019-08-20 report/chapter03.Rnw
final draft

colnames(cum.repr.trials.subg)  <- c("n","Number of groups", "Cumulative sum of groups")
#----------------------------------------------------------------------------------------------#

@@

\section{Structure and Variables}
The dataset consists of \Sexpr{format(length(unique(data$id)), big.mark = ",")} systematic reviews from the Cochrane Library with \Sexpr{format(length(unique(data$study.name)), big.mark = ",")} studies and \Sexpr{format(dim(data)[1], big.mark = ",")} results. A result of a study can be a primary or secondary or safety outcome, thus studies can contribute multiple results. The studies too a very large extent randomized and investigate the effects of healthcare and medical interventions and treatments. The reference to the treatment may be a placebo control or a different intervention or treatment. A result can not only be about efficacy of the treatments, but also about safety (adverse effects).\\
It will be continued with an example form the dataset. In Table \ref{barbiturate.row}, two results from a systematic review about barbiturates are shown as they are given in the dataset. As can be seen, further specifications are provided by the variables in the columns. \\
The \texttt{comparison.name} variable specifies \textit{what kind} of treatments or interventions are compared, the \texttt{outcome.name} variable \textit{how} it is compared, and the \texttt{subgroup.name} variable (not indicated in table) if and to \textit{what experimental subgroup} the result belongs to. \\
The result is of a binary type, and the counts of events in the treatment group are in \texttt{events1} and of the control group in \texttt{events2} and the total number of participants are given in columns \texttt{total1} and \texttt{total2}. As can be seen, events denote here "death at the end of follow-up".

@@

\begin{table}[ht]
  \begin{center}
    \begin{tabular}{l l}
      \textbf{Variable} & \textbf{Description}\\
      \hline
      \texttt{id} & An id of the review for identification purposes.\\
      \hline
      \texttt{study.name} & Name of the study to which the result belongs.\\
      \texttt{study.year} & Year in which the study was published.\\
      \hline
      \texttt{comparison.name/.nr} & Specification of the interventions compared in the study and a \\ &unique number for the comparison.\\
      \texttt{comparison.id} & Specification of the comparison by a string of the type ``CMP-xxx'' \\ &for the xxx. comparison within the review.\\
      \texttt{outcome.name/.nr} & Specification by which outcome the interventions are compared\\ &and a unique number for the outcome.\\
      \texttt{outcome.id} & Specification of the outcome by a string of the type \\ &``CMP-xxx.xx'' for the xx. outcome of the xxx. comparison \\ &within the review.\\
      \texttt{subgroup.name/.nr} & Potentially indication of affiliation to subgroups and a unique \\ &number for the subgroup.\\
      \texttt{subgroup.id} & Specification of the comparison by a string of the type \\ &``CMP-xxx.yy.xx'' for the xx. subgroup for the yy. outcome \\ & of the xxx. comparison within the review.\\
      \texttt{outcome.measure} & Outcome measure, \eg risk ratio.\\
      \texttt{outcome.measure.merged} & Outcome measure, \eg risk ratio, merged such that each method is uniquely classified.\\
      \texttt{outcome.flag} & A outcome flag to simplify programming; \\ &\texttt{DICH} for binary outcomes with fully available information \\ &on event counts,
      as given in a two-by-two \\ &table. \\ &\texttt{CONT} for continuous outcomes with available means and \\ &standard deviations. \\ & \texttt{IV} for all results with effects and standard errors but without \\ & the data necessary for their computation. \\ &\texttt{IPD} for individual patient data.\\
      \texttt{effect} & Magnitude of the effect given in the quantity denoted by \\ &\texttt{outcome measure}.\\
      \texttt{se} & Standard error of the measure of the effect.\\
      \texttt{events1/events2} & The counts of patients with an outcome if measurement/outcome \\ &is binary or dichotomous (1 for treatment group and 2 for control group).\\
      \texttt{total1/total2} & Number of patients in groups.\\
      \texttt{mean1/mean2} & Mean of patient measurements if outcome is continuous.\\
      \texttt{sd1/sd2} & Standard deviation if outcome is continuous.
    \end{tabular}
  \caption{Dataset variable names and descriptions  \label{variable}}
  \label{variable}
  \end{center}
\end{table}

\vspace{0mm}
The structure of a review will now be outlined based on an example of the dataset. The previously mentioned barbiturate and head injury review will be outlined. The aim was to ``assess the effects of barbiturates in reducing mortality, disability and raised ICP (intra-cranial pressure) in people with acute traumatic brain injury'' as well as to ``quantify any side effects resulting from the use of barbiturates'' %\citep{barbiturate}. \\
%Since there are arguments for and against use of barbiturates, the authors of the review did a comprehensive literature search and collected all available study findings. 
The review comprises five studies in total. Three of them compared barbiturate to placebo, one compared barbiturate to Mannitol and one Pentobarbital to Thiopental. The studies have different outcomes, for example, death or death and severe disability at follow up, but also dropout counts or adverse effects (secondary outcomes). 
We have continuous (e.g. mean body temperature) and binary outcome data (e.g. death/no death). One study split up outcomes for patients with and without haematoma, which would be subgroups. %It is important not to confuse results with studies. A study can contribute multiple results to a systematic review, for example, primary and secondary outcomes and adverse effects.

@@

The primary studies that are included in the reviews and have been published are most often from the years after 1980 (5\% quantile = \Sexpr{publication.year.range[1]}, 95\% quantile = \Sexpr{publication.year.range[6]},). The median of the publication years is \Sexpr{publication.year.range[3]}, the mean \Sexpr{round(publication.year.range[5])} and the quartiles are \Sexpr{publication.year.range[2]} and \Sexpr{publication.year.range[4]}. \Sexpr{format(publication.year.2018$n, big.mark = ",")} studies have been published in 2018.\\
The most frequent outcome measures are summarized in Table \ref{outcome.measure.frequencies}. One can conclude of the table that roughly 31 \% of outcomes in the dataset are continuous and the rest being some sort of discrete or binary outcomes, most often binary (more than 65\%).

@@

The sample sizes among results vary to some extent. There are 5\% of treatment group sample sizes that are smaller than \Sexpr{treatment.group.size.range[1]}, 95\% smaller than \Sexpr{treatment.group.size.range[6]}. The first quartile is \Sexpr{treatment.group.size.range[2]}, the median \Sexpr{treatment.group.size.range[3]}, the mean \Sexpr{round(treatment.group.size.range[5])} and the third quartile \Sexpr{treatment.group.size.range[4]}. The large difference between median and mean is caused by very large groups with over 2,000,000 participants. Analogously, the quantiles of the total sample size are: 5\% quantile = \Sexpr{samplesize.range[1]}, first quartile = \Sexpr{samplesize.range[2]}, median = \Sexpr{samplesize.range[3]}, third quartile = \Sexpr{samplesize.range[4]} and 95\% = \Sexpr{samplesize.range[6]}. The mean is \Sexpr{round(samplesize.range[5])}. \\
There are \Sexpr{comp.range[1]} reviews with five or fewer results. The 5\% and 95\% quantiles are \Sexpr{comp.range[7]} and \Sexpr{comp.range[6]}. The mean and median number of results per review are \Sexpr{study.range[5]} and \Sexpr{study.range[3]}, and the quartiles are \Sexpr{comp.range[2]} and \Sexpr{comp.range[4]}. Similarly, the number of reviews with a maximum of two studies included is \Sexpr{format(study.range[1], big.mark = ",")}, the mean study number is \Sexpr{study.range[5]}, the median \Sexpr{study.range[3]} and the interquartile range \Sexpr{study.range[2]} and \Sexpr{study.range[4]} and the 95\% quantile \Sexpr{study.range[6]}. The discrepancy between mean and median is due to large reviews with a high number of studies and results, most extreme in %\citealp{largest.review} 
which is a systematic review about antibiotic prophylaxis for preventing infection after cesarean section, with 95 studies and 1,497 results in total.\\
For results to be suitable for usage in meta-analysis, they have to be identical with respect to comparison and outcome. The studies in the dataset that have the same comparison, outcome and subgroup can be pooled in a meta-analysis. This distinction is also used by Cochrane, \ie the meta-analyses are identical to the meta-analyses done in the systematic reviews.\\ %Another approach that is not made here is to analyze all subgroups together, which is possible if enough studies are given.
The size of a meta-analysis denotes how many results are included in a group. Table \ref{repr.groups} shows the number of meta-analysis with size $\geq n$ results.

<<echo=FALSE, results='asis', warning =FALSE, message = FALSE>>=
rownames(cum.repr.trials.subg)[15] <- "$\\geq$ 15"
print(xtable(cum.repr.trials.subg, label = "repr.groups", caption = "Number and cumulative number of groups with meta-analysis size n.", align = "llrr", digits = 0), include.rownames = F, size = "footnotesize")
@

@@

The analysis includes all results with \texttt{outcome.flag} \texttt{DICH}, \texttt{CONT} and \texttt{IV}. If outcome is \texttt{IV}, the only meta-analyses with \texttt{outcome.measure.merged} = \texttt{OR} / \texttt{RR} / \texttt{SMD} / \texttt{MD} / \texttt{Hazard Ratio} / \texttt{Rate Ratio} are used. 
\citealp{Ioannidis2007} outlined criteria for application of small study effect tests:
\begin{itemize}
\item \textbf{Sample size}: A meta-analysis is comprised of at least ten studies ($n$ = \Sexpr{format(m.a.largerten.count, big.mark = ",")} remaining). 
\item \textbf{Study size}: The ratio between largest variance of an estimate and smallest variance of an estimate is larger than four ($n$ = \Sexpr{format(m.a.variance4.count, big.mark = ",")} remaining).
\item \textbf{Significance}: At least one treatment effect has a $p$-value below the significance threshold 0.05 ($n$ = \Sexpr{format(m.a.one.sig.count, big.mark = ",")} remaining)
\item \textbf{Heterogeneity}: The $I^2$ statistic of a given meta-analysis is smaller than 0.5, thus, the proportion of between study variance of the overall variance is smaller than 0.5 ($n$ = \Sexpr{format(m.a.smaller.5.count, big.mark = ",")} remaining).
\end{itemize}
Additionally, the following criteria have been applied:
\begin{itemize}
\item \textbf{Sensitivity Analyses}: When the same results are used multiple times for different meta-analyses, only one is retained. More precisely, if a study hat the same \texttt{study.name} and same \texttt{effect}, it was considered a duplicate, and the smaller meta-analysis of the two was excluded. The intention is to exclude sensitivity analyses which are operated on subsets of the available results.
\item \textbf{Zero events}: In the case of binary outcomes, meta-analyses with zero events in any study and any group are excluded ($n$ = \Sexpr{both.arms.zero.events.count} out of meta-analyses with at least ten studies).
\item \textbf{No withdrawn reviews}: Reviews that have been withdrawn are not included.
\item \textbf{No adverse effects}: No results of adverse effects of treatment are used for meta-analyses.we could thus clearly identify many safety results but this may be incomplete when safety outcomes are insufficiently declared, for example ``headache''.
\end{itemize}

>>>>

colnames(cum.repr.trials.subg)  <- c("n","", "cumulative number")
#----------------------------------------------------------------------------------------------#

@@

\section{Structure and Variables}
The dataset consists of \Sexpr{format(length(unique(data$id)), big.mark = ",")} systematic reviews from the Cochrane Library with \Sexpr{format(length(unique(data$study.name)), big.mark = ",")} studies and \Sexpr{format(dim(data)[1], big.mark = ",")} results. A result of a study can be a primary or secondary or safety outcome, thus studies can contribute multiple results. The studies too a very large extent randomized and investigate the effects of healthcare and medical interventions and treatments. The reference to the treatment may be a placebo control or a different intervention or treatment. A result can not only be about efficacy of the treatments, but also about safety (adverse effects).\\
It will be continued with an example form the dataset. In Table \ref{barbiturate.row}, two results from a systematic review about barbiturates are shown as they are given in the dataset. As can be seen, further specifications are provided by the variables in the columns. \\
The \texttt{comparison.name} variable specifies \textit{what kind} of treatments or interventions are compared, the \texttt{outcome.name} variable \textit{how} it is compared, and the \texttt{subgroup.name} variable (not indicated in table) if and to \textit{what experimental subgroups} the results belong to. \\
The result is of a binary type, and the counts of events in the treatment group are in \texttt{events1} and of the control group in \texttt{events2} and the total number of participants are given in columns \texttt{total1} and \texttt{total2}. As can be seen, events denote here "death at the end of follow-up".

@@

\begin{table}[ht]
  \begin{center}
    \begin{tabular}{ll}
      \textbf{Variable} & \textbf{Description}\\
      \hline
      \texttt{id} & An id of the review for identification purposes.\\
      \hline
      \texttt{study.name} & Name of the study to which the result belongs.\\
      \texttt{study.year} & Year in which the study was published.\\
      \hline
      \texttt{comparison.name/.nr} & Specification of the interventions compared in the study and a \\ &unique number for the comparison.\\
      \texttt{comparison.id} & Specification of the comparison by a string of the type ``CMP-xxx'' \\ &for the xxx. comparison within the review.\\
      \texttt{outcome.name/.nr} & Specification by which outcome the interventions are compared\\ &and a unique number for the outcome.\\
      \texttt{outcome.id} & Specification of the outcome by a string of the type \\ &``CMP-xxx.xx'' for the xx. outcome of the xxx. comparison \\ &within the review.\\
      \texttt{subgroup.name/.nr} & Potentially indication of affiliation to subgroups and a unique \\ &number for the subgroup.\\
      \texttt{subgroup.id} & Specification of the comparison by a string of the type \\ &``CMP-xxx.yy.xx'' for the xx. subgroup for the yy. outcome \\ & of the xxx. comparison within the review.\\
      \texttt{outcome.measure} & Outcome measure, for example risk ratio.\\
      \texttt{outcome.measure.merged} & Outcome measure, for example risk ratio, merged such \\ &that each method is uniquely classified.\\
      \texttt{outcome.flag} & A outcome flag to simplify programming; \\ &\texttt{DICH} for binary outcomes with fully available information \\ &on event counts,
      as given in a two-by-two \\ &table. \\ &\texttt{CONT} for continuous outcomes with available means and \\ &standard deviations. \\ & \texttt{IV} for all results with effects and standard errors but without \\ & the data necessary for their computation. \\ &\texttt{IPD} for individual patient data.\\
      \texttt{effect} & Magnitude of the effect given in the quantity denoted by \\ &\texttt{outcome measure}.\\
      \texttt{se} & Standard error of the measure of the effect.\\
      \texttt{events1/events2} & The counts of patients with an outcome if measurement/outcome \\ &is binary or dichotomous (1 for treatment group and 2 for control group).\\
      \texttt{total1/total2} & Number of patients in groups.\\
      \texttt{mean1/mean2} & Mean of patient measurements if outcome is continuous.\\
      \texttt{sd1/sd2} & Standard deviation if outcome is continuous.
    \end{tabular}
  \caption{Dataset variable names and descriptions  \label{variable}}
  \label{variable}
  \end{center}
\end{table}

\vspace{0mm}
The structure of a review will now be outlined based on an example of the dataset. The previously mentioned barbiturate and head injury review will be outlined. The aim was to ``assess the effects of barbiturates in reducing mortality, disability and raised ICP (intra-cranial pressure) in people with acute traumatic brain injury'' as well as to ``quantify any side effects resulting from the use of barbiturates'' %\citep{barbiturate}. \\
%Since there are arguments for and against use of barbiturates, the authors of the review did a comprehensive literature search and collected all available study findings.
The review comprises five studies in total. Three of them compared barbiturate treatment to placebo, one compared barbiturate to Mannitol and one Pentobarbital to Thiopental. The studies have different outcomes, for example, death or death and severe disability at follow up, but also dropout counts or adverse effects (secondary outcomes).
We have continuous (e.g. mean body temperature) and binary outcome data (e.g. death/no death). One study split up outcomes for patients with and without haematoma, which would be subgroups. %It is important not to confuse results with studies. A study can contribute multiple results to a systematic review, for example, primary and secondary outcomes and adverse effects.

@@

The primary studies that are included in the reviews and have been published are most often from the years after 1980 (5\% quantile = \Sexpr{publication.year.range[1]}, 95\% quantile = \Sexpr{publication.year.range[6]},). The median of the publication years is \Sexpr{publication.year.range[3]}, the mean \Sexpr{round(publication.year.range[5])} and the quartiles are \Sexpr{publication.year.range[2]} and \Sexpr{publication.year.range[4]}. \Sexpr{format(publication.year.2018$n, big.mark = ",")} studies have been published in 2018.\\
The most frequent outcome measures are summarized in Table \ref{outcome.measure.frequencies}. One can conclude of the table that roughly 31\% of outcomes in the dataset are continuous and the rest being some sort of discrete or binary outcomes, most often binary (more than 65\%).

@@

The sample sizes among results vary to some extent. There are 5\% of treatment group sample sizes that are smaller than \Sexpr{treatment.group.size.range[1]}, 95\% smaller than \Sexpr{treatment.group.size.range[6]}. The first quartile is \Sexpr{treatment.group.size.range[2]}, the median \Sexpr{treatment.group.size.range[3]}, the mean \Sexpr{round(treatment.group.size.range[5])} and the third quartile \Sexpr{treatment.group.size.range[4]}. The large difference between median and mean is caused by very large groups with over 2,000,000 participants. Analogously, the quantiles of the total sample size are: 5\% quantile = \Sexpr{samplesize.range[1]}, first quartile = \Sexpr{samplesize.range[2]}, median = \Sexpr{samplesize.range[3]}, third quartile = \Sexpr{samplesize.range[4]} and 95\% = \Sexpr{samplesize.range[6]}. The mean is \Sexpr{round(samplesize.range[5])}. \\
There are \Sexpr{comp.range[1]} reviews with five or fewer results. The 5\% and 95\% quantiles are \Sexpr{comp.range[7]} and \Sexpr{comp.range[6]}. The mean and median number of results per review are \Sexpr{study.range[5]} and \Sexpr{study.range[3]}, and the quartiles are \Sexpr{comp.range[2]} and \Sexpr{comp.range[4]}. Similarly, the number of reviews with a maximum of two studies included is \Sexpr{format(study.range[1], big.mark = ",")}, the mean study number is \Sexpr{study.range[5]}, the median \Sexpr{study.range[3]} and the interquartile range \Sexpr{study.range[2]} and \Sexpr{study.range[4]} and the 95\% quantile \Sexpr{study.range[6]}. The discrepancy between mean and median is due to large reviews with a high number of studies and results, most extreme in %\citealp{largest.review}
which is a systematic review about antibiotic prophylaxis for preventing infection after cesarean section, with 95 studies and 1,497 results in total.\\
For results to be suitable for usage in meta-analysis, they have to be identical with respect to comparison and outcome. The studies in the dataset that have the same comparison, outcome and subgroup can be pooled in a meta-analysis. This distinction is also used by Cochrane, \ie the meta-analyses are identical to the meta-analyses done in the systematic reviews.\\ %Another approach that is not made here is to analyze all subgroups together, which is possible if enough studies are given.
% The size of a meta-analysis denotes how many results are included in a group. Table \ref{repr.groups} shows the number of meta-analysis with size $\geq n$ results.

% 
% <<echo = FALSE, results = 'asis'>>=
% print(xtable(cum.repr.trials.subg[,-2], label = "repr.groups", caption = "Cumulative number of meta-analyses with $n$ or more results.", align = "llr", digits = 0), include.rownames = F,  size = "footnotesize", floating = TRUE)
% @
%

% 
% <<echo=FALSE, results='asis', warning =FALSE, message = FALSE>>=
% # rownames(cum.repr.trials.subg)[15] <- "$\\geq$ 15"
% cum.repr.trials.subg <- cum.repr.trials.subg[,-2]
% print(xtable(cum.repr.trials.subg[,-2], label = "repr.groups", caption = "Cumulative number of groups with n results.", align = "llr", digits = 0), include.rownames = F,  size = "footnotesize", sanitize.rownames.function = identity)
% @

@@

The analysis includes all results with \texttt{outcome.flag} \texttt{DICH}, \texttt{CONT} and \texttt{IV}. If outcome is \texttt{IV}, the only meta-analyses with \texttt{outcome.measure.merged} = \texttt{OR} / \texttt{RR} / \texttt{SMD} / \texttt{MD} / \texttt{Hazard Ratio} / \texttt{Rate Ratio} are used. 
\citealp{Ioannidis2007} outlined criteria for application of small study effect tests:
\begin{itemize}
\item \textbf{Sample size}: A meta-analysis is comprised of at least ten studies ($n$ = \Sexpr{format(m.a.largerten.count, big.mark = ",")} remaining). 
\item \textbf{Study size}: The ratio between largest variance of an estimate and smallest variance of an estimate is larger than four ($n$ = \Sexpr{format(m.a.variance4.count$n, big.mark = ",")} remaining).
\item \textbf{Significance}: At least one treatment effect has a $p$-value below the significance threshold 0.05 ($n$ = \Sexpr{format(m.a.one.sig.count$n, big.mark = ",")} remaining)
\item \textbf{Heterogeneity}: The $I^2$ statistic of a given meta-analysis is smaller than 0.5, thus, the proportion of between study variance of the overall variance is smaller than 0.5 ($n$ = \Sexpr{format(m.a.smaller.5.count$nn, big.mark = ",")} remaining).
\end{itemize}
Additionally, the following criteria have been applied:
\begin{itemize}
\item \textbf{Sensitivity Analyses}: When the same results are used multiple times for different meta-analyses, only one is retained. More precisely, if a study hat the same \texttt{study.name} and same \texttt{effect}, it was considered a duplicate, and the smaller meta-analysis of the two was excluded. The intention is to exclude sensitivity analyses which are operated on subsets of the available results.
\item \textbf{Zero events}: In the case of binary outcomes, meta-analyses with zero events in any study and any group are excluded ($n$ = \Sexpr{both.arms.zero.events.count} out of meta-analyses with at least ten studies).
\item \textbf{No withdrawn reviews}: Reviews that have been withdrawn are not included.
\item \textbf{No adverse effects}: No results of adverse effects of treatment are used for meta-analysis. We could thus clearly identify many safety results but this may be incomplete when safety outcomes are insufficiently declared, for example ``headache''.
\end{itemize}
